Network Deliberation: The role of network structure in large-scale, internet-enabled, participatory decision-making

### 2022-05-05 UMich/chapters/ch05-discussion.tex
Edits.

\section{Conclusion}
Inexpensive, instantaneous, global, peer-to-peer communication has only been possible for a tiny sliver of recent human history.
Existing organizational and institutional structures have evolved over hundreds or thousands of years.
While the full implications of digital communication for human institutions can't be predicted, our institutions have already started adapting in exciting ways.
In particular, we have seen the emergence of very large collaborations with participatory governance, which would have seemed impossible before the internet.
In the research presented here, I have empirically identified one of the commonalities between these unprecedented projects, namely interlocking networks of small pods: network deliberation.
The agent-based simulation and experiment presented here study the properties of a simplified model of network deliberation, in order to better unerstand the principles and mechanisms at work.
This research contributes knowledge on the interaction of network deliberaiton with factors such as network efficiency, individual behavior, and in particiular, social influence.
It is my hope that this work will help enable large groups to collaborate on common goals in a manner that allows all members to contribute and allows all voices to be heard.

>>>>

\section{Conclusion}
Inexpensive, instantaneous, global, peer-to-peer communication has only been possible for a tiny sliver of recent human history.
Existing organizational and institutional structures have evolved over hundreds or thousands of years.
While the full implications of digital communication for human institutions can't be predicted, our institutions have already started adapting in exciting ways.
In particular, we have seen the emergence of very large collaborations with participatory governance, which would have seemed impossible before the internet.
In the research presented here, I have empirically identified one of the commonalities between these unprecedented projects, namely interlocking networks of small pods: network deliberation.
The agent-based simulation and experiment presented here study the properties of a simplified model of network deliberation, in order to better unerstand the principles and mechanisms at work.
This research contributes knowledge on the interaction of network deliberaiton with factors such as network efficiency, individual behavior, and in particiular, social influence.
It is my hope that this work will help large groups collaborate on common goals in a manner that allows all members to contribute and allows all voices to be heard.

### 2022-05-05 UMich/front_materials/abstract.tex
Edits.

As groups grow in size, they gain access to additional resources,
creating opportunities for collective intelligence and collective action.
However, at very large scales, group decision-making becomes prohibitively slow and difficult
to coordinate.
Traditional solutions include representative decision-making and/or a shift from deliberation
to voting.
Both approaches sacrifice desirable properties.
Representative decision-making loses the potential benefits of collective intelligence and
introduces hierarchies that may place the interests of specific individuals ahead of the
interests of the group.
Voting sacrifices generativity: allowing a choice between predefined options,
without allowing for improvement to those options.
Arrow’s impossibility theorem also fundamentally limits the fairness of voting.
This project proposes Networked Deliberation as a potential means of large-scale decision-making.
In Networked Deliberation, members of a large group are repeatedly partitioned into small
deliberative pods.
Overlap between pods at different stages enables group-wide diffusion of information and
preferences.
Different methods for assigning members to pods result in different network topologies.
In partnership with large community groups, this project will use a web-based platform to
evaluate Network Deliberation.
For a variety of network topologies, the speed and quality of preference convergence will be
evaluated using quantitative information from ranked-choice polls throughout the process.
Deliberator satisfaction and experience will also be evaluated using qualitative analysis of
deliberative text and surveys.
This work seeks to improve the ability of very large groups to quickly develop a consensus,
enabling a more effective use of shared resources to achieve common goals.

>>>>

As groups grow in size, they gain access to additional resources, creating opportunities for collective intelligence and collective action.
However, at very large scales, group decision-making becomes prohibitively slow and difficult to coordinate.
Traditional solutions include representative decision-making and/or a shift from deliberation to voting.
Both approaches sacrifice desirable properties.
Representative decision-making loses the potential benefits of collective intelligence and introduces hierarchies that may place the interests of specific individuals ahead of the interests of the group.
Voting sacrifices generativity: allowing a choice between predefined options, without allowing for improvement to those options.
Arrow’s impossibility theorem also fundamentally limits the fairness of voting.
This project proposes Networked Deliberation as a potential means of effective large-scale decision-making.
In Networked Deliberation, members of a large group are repeatedly partitioned into small deliberative pods.
Overlap between pods at different stages enables group-wide diffusion of information and preferences.
Different methods for assigning members to pods result in different network topologies.
This work combines observational study, agent-based modeling, and experiment to evaluate and better understand network deliberation.
An empirical observation of WikiProjects on the English-language Wikipeida identifies the network properties of the most effective projects.
A simple agent-based model of Network Deliberation shows improvements over conventional deliberation in the presence of strong social influence.
Finally, a controlled experiment studies network deliberation in a real world setting, tracking how individual preferences evolve, and finding evidence that network deliberation provides protection against negative consequences of social influence.
This work seeks to improve the ability of very large groups to quickly develop a consensus, enabling a more effective use of shared resources to achieve common goals.
